\chapter{Abstract}

The field of automatic Text Categorization (TC) concerns the creation
of categorizer functions, usually involving Machine Learning
techniques, to assign labels from a pre-defined set of categories to
documents based on the documents' content.  Because of the many
variations on how this can be achieved and the diversity of
applications in which it can be employed, creating specific TC
applications is often a difficult task.

This thesis concerns the design, implementation, and testing of an
Object-Oriented Application Framework for Text Categorization.  By
encoding expertise in the architecture of the framework, many of the
barriers to creating TC applications are eliminated.  Developers can
focus on the domain-specific aspects of their applications, leaving
the generic aspects of categorization to the framework.  This allows
significant code and design reuse when building new applications.

Chapter \ref{intro} provides an introduction to automatic Text
Categorization, Ob\-ject-Oriented Application Frameworks, and Design
Patterns.  Some common application areas and benefits of using
automatic TC are discussed.  Frameworks are defined and their
advantages compared to other software engineering strategies are
presented.  Design patterns are defined and placed in the context of
framework development.  An overview of three related products in the
TC space, Weka, Autonomy, and Teragram, follows.

Chapter \ref{background-tc} contains a detailed presentation of Text
Categorization.  TC is formally defined, followed by a detailed
account of the main functional areas in Text Categorization that a
modern TC framework must provide.  These include document tokenizing,
feature selection and reduction, Machine Learning techniques, and
categorization runtime behavior.  Four Machine Learning techniques
(\naive\ Bayes categorizers, k-Nearest-Neighbor categorizers, Support
Vector Machines, and Decision Trees) are presented, with discussions
of their core algorithms and the computational complexity involved.
Several measures for evaluating the quality of a categorizer are then
defined, including precision, recall, and the $F_\beta$ measure.

The design of a framework that addresses the functional areas from
Chapter \ref{background-tc} is presented in Chapter \ref{design}.
This design is motivated by consideration of the framework's audience
and some expected usage scenarios.  The core architectural classes in
the framework are then presented, and Design Patterns are employed in
a detailed discussion of the cooperative relationships among framework
classes.  This is the first known use of Design Patterns in an
academic work on Text Categorization software.  Following the
presentation of the framework design, some possible design limitations
are discussed.

The design in Chapter \ref{design} has been implemented as the \aicat\
Perl package.  Chapter \ref{Implementation} is a short discussion of
implementation issues, including considerations in choosing the
programming language.  Special consideration is given to the
implementation of constructor methods in the framework, since they are
responsible for enforcing the structural relationships among framework
classes.  Three data structure issues within the framework are then
discussed: feature vectors, sets of document or category objects, and
the serialized representation of a framework object.

Chapter \ref{Evaluation} evaluates the framework from several
different perspectives on two corpora.  The first corpus is the
standard Reuters-21578 benchmark corpus, and the second is assembled
from messages sent to an educational ask-an-expert service.  Using
these corpora, the framework is evaluated on the measures introduced
in Chapter \ref{background-tc}.  The performance on the first corpus
is compared to the well-known results in \cite{yang:99}.  The \naive\
Bayes categorizer is found to be competitive with standard
implementations in the literature, and the Support Vector Machine and
k-Nearest-Neighbor implementations are outperformed by comparable
systems by other researchers.  The framework is then evaluated in
terms of its resource usage, and several applications using \aicat\
are presented in order to show the framework's ability to function in
the usage scenarios discussed in Chapter \ref{design}.


\chapter{Acknowledgments}

I would like to thank Rafael Calvo for his expert supervision of my
thesis, and for giving me the opportunity to pursue this project.  The
rest of the Web Engineering Group---Jae-Moon Lee, Xiaobo Li, Nick
Carroll, and Gosia Mandrela---provided valuable testing and feedback
on \aicat\ and created a quite pleasant research environment.  The
Language Technology group---particularly Casey Whitelaw, Elisabeth
Crawford, and Jon Patrick---was also a good source of feedback and
inspiration.  The Open-Source community provided incentives to write
clean, usable, documented software by their mere existence.  Sheri
Schechinger proofread this document and helped make countless
readability improvements.  Research
Assistantship funding from the Capital Markets Co-operative Research
Centre provided much-appreciated support for research on financial
corpora.

On a more personal level, I would like to thank Sheri Schechinger for sticking
with me across ten thousand miles of ocean, and the city of Sydney for
being such a great place to spend time.

\chapter{Preface}

This thesis is the culmination of a Masters project in the Web
Engineering Group at the University of Sydney School of Electrical and
Information Engineering.  The project has produced two large
products---one is this thesis, and the other is the \aicat\ framework
itself, which forms the subject matter of most of the thesis.

In order to produce such a framework, research into current Text
Categorization algorithms has been necessary, as well as research into
software engineering practices for building object-oriented
frameworks.  The discourse in this thesis does not assume any prior
familiarity with Text Categorization, but it does assume that the
reader is familiar with the basic concepts and terms of
object-oriented programming, such as ``class,'' ``object,'' and
``instance.''


\section*{Availability}

The latest released version of the \aicat\ framework (currently 0.04)
is always available at
\url{http://search.cpan.org/author/KWILLIAMS/}.  Perl source code,
documentation, and a simple example application are included in the
distribution.

For developers who wish to stay more actively involved with tracking
changes in the framework, the entire distribution is also available
using the Concurrent Versions System (CVS).  This allows developers to
access the latest bug fixes, to create their own patches against the
main framework code, and to track changes between releases.  Details
of how to access the CVS version are at
\url{http://sourceforge.net/cvs/?group_id=62831}, or via the project's
development home page at
\url{http://sourceforge.net/projects/ai-categorizer/}.

The ApteMod data set discussed in Chapter \ref{Evaluation} is
available for download from
\url{http://kdd.ics.uci.edu/databases/reuters21578/reuters21578.html}.

The Dr. Math data set is not available for direct download, but
interested parties may contact Ken Williams at \url{ken@mathforum.org}
for details.

After submission to the University of Sydney, this thesis document
will be available in electronic format at
\url{http://www.ee.usyd.edu.au/~kenw/Thesis.pdf}, and in hardcopy
format from the University of Sydney Engineering Library.

\section*{Licensing}

The \aicat\ framework is implemented as a set of Perl modules (see
Section \ref{imp-language}).  As is customary with many Perl modules,
the framework is distributed under the same licensing terms as the
standard version of the Perl interpreter.  This means that the user
may choose either the GNU General Public License or the Artistic
License as the terms of using the software, whichever fits better with
their needs.  In practical terms, this means that the code is
encouraged to be used in research, commercial, educational, or other
environments, without the need to pay royalties to the software's
original author.  It also means that the software's inner workings are
available to be inspected or modified by other developers for their
own projects.

Licenses of the above type are called ``open source'' licenses.  Their
goal is to foster the development and evolution of software by
leveraging the user community and developer community as a resource
that can feed back into the development cycle.  According to
\url{http://www.opensource.org/}, ``open source promotes software
reliability and quality by supporting independent peer review and
rapid evolution of source code.''  This aligns very well with the
traditional goals of academic research.  By making the source code
discussed in academic publications available as open source resources, the
results can far more easily be verified by other researchers.

For more information on open source concepts, please visit
\url{http://www.opensource.org/}.

This thesis is copyright \copyright 2003 by Ken Williams. This
material may be distributed only subject to the terms and conditions
set forth in the Open Publication License, v1.0 or later (the latest
version is presently available at
\url{http://www.opencontent.org/openpub/}).

\section*{Motivations}

My own personal motivations for embarking on this project were to
further educate myself on Text Categorization research, to learn more
about framework methodology, and to provide a software resource to
others who wish to use TC methods in their software projects.  I have
long been interested in Machine Learning methods for various purposes,
and I enjoy working with natural languages.  Doing corpus-based work
in Natural Language Processing is a fun combination of Machine
Learning and Linguistics, and reading the literature on the topic
always makes me excited to work on my next project.

Unfortunately, when I was just beginning to do work on my own Text
Categorization projects, I found that there were very few TC tools
freely available for my use, and those that were available were often
difficult to customize.  Very few tools were available in Perl, my
usual language of choice, and this seemed like an odd situation given
the well-known agility of Perl at handling text data.  I cannot hope
to solve everyone's software needs in TC, but the \aicat\ framework
represents my best effort in providing the kind of thing I was looking
for when I began working in this area.

My interest in framework development has recently increased by working
on the \class{HTML::Mason} project \cite{rolsky:02}.  For this, I
and others helped shepherd the code from a fairly monolithic
function-based tool to a customizable OO framework suitable for many
more purposes than it was originally developed to serve.  I became
convinced of the power of framework development with that project, and
I sought to bring the same benefits to an open-source framework for TC.


\section*{Contributions}

During the course of the candidature on which this thesis is based,
the following contributions were accomplished:

\begin{itemize}
\item The \aicat\ framework was designed, implemented, and released
  under an open-source license \cite{cpan}.  The release includes
  documentation and a simple example application using the framework.
\item \naive\ Bayes and Decision Tree categorizers were implemented,
  as well as a mechanism which allows users to use categorizers
  implemented in the Weka Machine Learning system \cite{weka:99}.  A
  simple probabilistic guessing categorizer has also been implemented
  to provide a baseline for experimentation.
\item The framework currently has a Document Frequency feature
  selection module implemented.
\item A paper on the design and applicability of the \aicat\ framework
  was published in the proceedings of the 7th Australasian Document
  Computing Symposium \cite{williams:02}.
\item A short paper on the use of the \aicat\ framework to categorize
  financial documents was published in the proceedings of the
  7th Australasian Document Computing Symposium \cite{calvo:02}.
\item A paper on the use of \aicat\ to automatically categorize
  mathematics questions will be published in the
  proceedings of the 11th International Conference on Artificial Intelligence in
  Education \cite{williams:03}.
\item New testing corpora have been assembled in the educational and
  financial domains, and the framework has been evaluated using them
  (see Chapter \ref{Evaluation}).
\item Contributions from other developers have provided the framework
  with an SVM categorizer.  Collaborative work with other developers
  have provided Rocchio and k-Nearest-Neighbor categorizers.
\item An overview seminar on TC and the design of \aicat\ was given at
  the University of Sydney.  An invited presentation of the same
  seminar was given to the Language Technology group at Macquarie
  University.
\item Tutorials on Machine Learning were presented at the O'Reilly
  2002 Open Source Conference and 2003 Bioinformatics Technology
  Conference (\url{http://conferences.oreilly.com/}).
\end{itemize}

